---
title: "SEM 3"
author: "Simon Roth"
date: "04.05.2017"
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
  html_document:
    theme: spacelab
  rmarkdown::github_document: default
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{rotating}
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = F # Whether to display code along with its results
                      , eval = T # Whether to evaluate the code and include its results
                      , results = "hide" # this at deafult is in end much more efficient
                      , cache = F # Whether to cache results for future renders (efficient!)
                      , warning = F # Whether to display errors
                      , message = F # Whether to display messages
                      , error = F # mybe turn on
                      , tidy = F # Whether to reformat code in a tidy way when displaying it
                      #, root.dir = normalizePath('/Users/simonroth/Dropbox/methods/git/tidy_textmining/my_summary/') 
                      #, root.dir = normalizePath("E:/Dropbox/methods/git/tidy_textmining")
                      # in order to calculate relative paths
                      , fig.width = 6
                      , fig.height = 4
                      , fig.align = "center"
                      )
```

```{r load, echo=F}
rm(list=ls())
load("/Users/simonroth/Dropbox/uni/MA/4_semester/SEM/data/allbus_final/allbus_Rdata.Rdata")
```


## Itembatterien 

**Islamophobie:** 

* mm01: ISLAMAUSUEBUNG IN BRD BESCHRAENKEN
    + -10 Befragter gehört einer islamischen Religionsgemeinschaft an (Code 1 in rd03)
    + -9
    + 1 Stimme überhaupt nicht zu
    + 2
    + 3
    + 4
    + 5 
    + 6 
    + 7 Stimme voll und ganz zu
* mm02: ISLAM PASST IN DIE DEUTSCHE GESELLSCHAFT
* mm03: ANWESENHEIT VON MUSLIMEN BRINGT KONFLIKT
* mm04: STAAT SOLLTE ISLAM. GRUPPEN BEOBACHTEN
* mm05: MUSLIMISCHER BUERGERMEISTER IN ORDNUNG
* mm06: UNTER MUSLIMEN SIND VIELE REL. FANATIKER



**Nationalbewusstsein:**

* mn11: DEUTSCH SEIN: DEUTSCHE STAATSBUERGERSCH.
* mn12: DEUTSCH SEIN: CHRISTL.RELIGION ZUGEHOER.
* mn13: DEUTSCH SEIN: BEKENNTNIS ZUR DEMOKRATIE
* mn14: DEUTSCH SEIN: VIELE DEUTSCHE BEKANNTE
* mn15: DEUTSCH SEIN: ALTE STAATSANGEH.AUFGEBEN
* mn16: DEUTSCH SEIN: VERBUNDENHEIT ZU DEUTSCHL.
* mn17: DEUTSCH SEIN: ALTE GEBRAEUCHE ABLEGEN
* mn18: DEUTSCH SEIN: GUT DEUTSCH SPRECHEN
* mn19: DEUTSCH SEIN: WESTLICHE WERTE TEILEN
* mn20: DEUTSCH SEIN: MIND. 1 ELTERNTEIL DEUTSCH
* mn21: DEUTSCH SEIN: IN DEUTSCHLAND GEBOREN

**Sozio-Ökonomische Variablen**

* sex: GESCHLECHT (Int.: Geschlecht der befragten Person ohne Befragen eintragen!)
    + 1 Männlich 
    + 2 Weiblich
* age: ALTER: metrisch
* agec: ALTER: KATEGORISIERT 6
    + 18 - 29 Jahre
    + 30 - 44 Jahre
    + 45 - 59 Jahre
    + 60 - 74 Jahre
    + 75 - 89 Jahre
    + Über 89 Jahre
* isced97: BEFR.: ISCED 1997 - 6 STUFEN: International Standard Classification of Education (ISCED) 1997, 6 Stufen
    1. Level - Primary education or first stage of basic education
    2. Level - Lower secondary or second stage of basic education
    3. Level - (Upper) secondary education
    4. Level - Post-secondary non-tertiary education
    5. Level - First stage of tertiary education
    6. Level - Second stage of tertiary education
* Allgemeiner Bildungsabschluss?


```{r}
###--------------------------------------------------------------------------###
### Listewise Deletion
###--------------------------------------------------------------------------###

library(dplyr)
#glimpse(dat)

### Nationalbewusstsein
nation_ld <- dat %>%
  select(ger_citiz:born_ger) %>%
  filter(dat$sp_nation == 0) %>%
  na.omit()

nation_ld_sum <- sapply(nation_ld, 
  plyr::each(mean, median, sd, var, min, max)) %>%
  t() %>%
  as.data.frame()

nation_ld_sum$method <- "Listwise deletion"


### Islamophobie
islam_ld <- dat %>%
  select(stop_islam:muslim_rad) %>%
  na.omit()

islam_ld_sum <- sapply(islam_ld, 
  plyr::each(mean, median, sd, var, min, max)) %>%
  t() %>%
  as.data.frame()

islam_ld_sum$method <- "Listwise deletion"
```



```{r}
###--------------------------------------------------------------------------###
### Mean Imputation 1 (hand-made)
###--------------------------------------------------------------------------###

### Nationalbewusstsein
nation_mean1 <- dat %>% 
  select(ger_citiz:born_ger) %>% 
  filter(dat$sp_nation == 0)

for(i in 1:length(nation_mean1)){
  nation_mean1[is.na(nation_mean1[[i]]), i] <- 
    mean(nation_mean1[[i]], na.rm = T)
}

nation_mean1_sum <- sapply(nation_mean1, 
  plyr::each(mean, median, sd, var, min, max)) %>%
  t() %>%
  as.data.frame()

nation_mean1_sum$method <- "Unconditional mean imputation"

### Islamophobie
islam_mean1 <- dat %>% 
  select(stop_islam:muslim_rad) %>% 
  filter(dat$sp_islam == 0)

for(i in 1:length(islam_mean1)){
  islam_mean1[is.na(islam_mean1[[i]]), i] <- 
    mean(islam_mean1[[i]], na.rm = T)
}

islam_mean1_sum <- sapply(islam_mean1, 
  plyr::each(mean, median, sd, var, min, max)) %>%
  t() %>%
  as.data.frame()

islam_mean1_sum$method <- "Unconditional mean imputation"
```



# MICE

MICE can handle both MAR and Missing Not at Random (MNAR) (p. 15).

**Options for method:**

* `pmm` 			Predictive mean matching (any)
* `norm` 		Bayesian linear regression (numeric)
* `norm.nob` 		Linear regression ignoring model error (numeric)
* `norm.boot` 		Linear regression using bootstrap (numeric)
* `norm.predict`	Linear regression, predicted values (numeric)
* `mean` 		Unconditional mean imputation (numeric)
* `2l.norm` 		Two-level normal imputation (numeric)
* `2l.pan` 		Two-level normal imputation using pan (numeric)
* `2lonly.mean` 	Imputation at level-2 of the class mean (numeric)
* `2lonly.norm` 	Imputation at level-2 by Bayesian linear regression (numeric)
* `2lonly.pmm`		Imputation at level-2 by Predictive mean matching (any)
* `quadratic` 		Imputation of quadratic terms (numeric)
* `logreg` 		Logistic regression (factor, 2 levels)
* `logreg.boot` 	Logistic regression with bootstrap
* `polyreg` 		Polytomous logistic regression (factor, >= 2 levels)
* `polr`			Proportional odds model (ordered, >=2 levels)
* `lda` 			Linear discriminant analysis (factor, >= 2 categories)
* `cart`			Classification and regression trees (any)
* `rf` 			Random forest imputations (any)
* `ri` 			Random indicator method for nonignorable data (numeric)
* `sample` 		Random sample from the observed values (any)
* `fastpmm` 		Experimental: Fast predictive mean matching using C++ (any)

The functions mice.impute.norm() and mice.impute.norm.nob() impute according
to a linear imputation model, and are fast and efficient if the model residuals are close to normal. The second model ignores any sampling uncertainty of the imputation model, so it is only appropriate for very large samples



```{r, eval = T}
###--------------------------------------------------------------------------###
### Mean Imputation 2 (mice)
###--------------------------------------------------------------------------###
## Unconditional mean imputation (numeric)
library(mice)
### Nationalbewusstsein

nation_mean2 <- dat %>% 
  select(ger_citiz:born_ger)  %>%
  filter(dat$sp_nation == 0)

nation_mean2_imp <- mice(nation_mean2, 
          method = rep("mean", 11))
nation_mean2_imp_final <- complete(nation_mean2_imp)

### summerize data
nation_mean2_imp_final <- nation_mean2_imp_final %>% 
  select(ger_citiz:born_ger)

nation_mean2_imp_sum <- sapply(nation_mean2_imp_final, 
  plyr::each(mean, median, sd, var, min, max)) %>%
  t() %>%
  as.data.frame() %>%
  mutate(method = "Unconditional mean imputation")


### Islamophobie

islam_mean2 <- dat %>% 
  select(stop_islam:muslim_rad) %>% 
  filter(dat$sp_islam == 0)

islam_mean2_imp <- mice(islam_mean2, 
          method = c("mean", "mean", "mean", "mean", "mean", "mean"))
islam_mean2_imp_final <- complete(islam_mean2_imp)

### summerize data
islam_mean2_imp_final <- islam_mean2_imp_final %>% 
  select(stop_islam:muslim_rad)

islam_mean2_imp_sum <- sapply(islam_mean2_imp_final, 
  plyr::each(mean, median, sd, var, min, max)) %>%
  t() %>%
  as.data.frame() %>%
  mutate(method = "Unconditional mean imputation")
```



```{r, result = "asis"}
###--------------------------------------------------------------------------###
### Linear regression, predicted values (numeric; mice)
###--------------------------------------------------------------------------###

### Islamophobie

# select data for imp islamophobia
islam_lm <- dat %>% 
  select(respid:age, isced97, stop_islam:muslim_rad)

# get initial mice object
ini <- mice(islam_lm, maxit = 0, print = F)

attributes(ini)

# default methods
method <- ini$method
# normal regression
method <- c(rep("", 4) ,rep("norm.predict", 6))

# postprocessing after imputation run
post<-ini$post
# avoid negative imputed values
post <- c(rep("", 4) ,rep("imp[[j]][,i]<-squeeze(imp[[j]][,i],c(0,7))", 6))

### run linear imputation
islam_lm_imp <- mice(islam_lm ,
              method = method,
              post = post,
              seed = 1234)

# plot(islam_lm_imp)

islam_lm_imp_final <- complete(islam_lm_imp)
table(is.na(islam_lm_imp_final$muslim_surv))

### summerize data
islam_lm_imp_final <- islam_lm_imp_final %>% 
  select(stop_islam:muslim_rad)

islam_lm_imp_sum <- sapply(islam_lm_imp_final, 
  plyr::each(mean, median, sd, var, min, max)) %>%
  t() %>%
  as.data.frame() %>%
  mutate(method = "Linear regression, predicted values")
```



```{r}
###--------------------------------------------------------------------------###
### Linear regression, predicted values (numeric; mice)
###--------------------------------------------------------------------------###

### Nationalbewusstsein

nation_lm <- dat %>%
  select(respid:age, isced97:born_ger) %>%
  filter(dat$sp_nation == 0)


### initialize impuation object
# get initial mice object
ini <- mice(nation_lm, maxit = 0, print = F)
attributes(ini)


### default settings
# default methods
method <- ini$method
# normal regression
method <- c(rep("", 4) ,rep("norm.predict", 11))

# postprocessing after imputation run
post<-ini$post
# avoid negative imputed values
post <- c(rep("", 4) ,rep("imp[[j]][,i]<-squeeze(imp[[j]][,i],c(0,7))", 11))


### run linear imputation
nation_lm_imp <- mice(nation_lm ,
              method = method,
              post = post,
              seed = 1234)

# plot(nation_lm_imp)

nation_lm_imp_final <- complete(nation_lm_imp)
table(is.na(nation_lm_imp_final$x))

### summerize data
nation_lm_imp_final <- nation_lm_imp_final %>% 
  select(ger_citiz:born_ger)

nation_lm_imp_sum <- sapply(nation_lm_imp_final, 
  plyr::each(mean, median, sd, var, min, max)) %>%
  t() %>%
  as.data.frame() %>%
  mutate(method = "Linear regression, predicted values")
```



```{r}
###--------------------------------------------------------------------------###
### Predictive mean matching (any)
###--------------------------------------------------------------------------###

### Nationalbewusstsein
# select data for imp 
nation_pmm <- dat %>% 
  select(respid:age, isced97, ger_citiz:born_ger)

# default methods
method <- c(rep("", 4) ,rep("pmm", 11))
# avoid negative imputed values
post <- c(rep("", 4) ,rep("imp[[j]][,i]<-squeeze(imp[[j]][,i],c(0,7))", 11))

### run linear imputation
nation_pmm_imp <- mice(nation_pmm ,
              method = method,
              post = post,
              seed = 1234)

# plot(islam_lm_imp)

nation_pmm_imp_final <- complete(nation_pmm_imp)
table(is.na(nation_pmm_imp_final$muslim_surv))

### summerize data
nation_pmm_imp_final <- nation_pmm_imp_final %>% 
  select(ger_citiz:born_ger)

nation_pmm_imp_sum <- sapply(nation_pmm_imp_final, 
  plyr::each(mean, median, sd, var, min, max)) %>%
  t() %>%
  as.data.frame() %>%
  mutate(method = "Predictive mean matching")
```



```{r}
###--------------------------------------------------------------------------###
### Predictive mean matching (any)
###--------------------------------------------------------------------------###

### Islamophobie
# select data for imp 
islam_pmm <- dat %>% 
  select(respid:age, isced97, stop_islam:muslim_rad)

# default methods
method <- c(rep("", 4) ,rep("pmm", 6))
# avoid negative imputed values
post <- c(rep("", 4) ,rep("imp[[j]][,i]<-squeeze(imp[[j]][,i],c(0,7))", 6))

### run linear imputation
islam_pmm_imp <- mice(islam_pmm,
              method = method,
              post = post,
              seed = 1234)

# plot(islam_lm_imp)

islam_pmm_imp_final <- complete(islam_pmm_imp)
#table(is.na(islam_lm_imp_final$muslim_surv))

### summerize data
islam_pmm_imp_final <- islam_pmm_imp_final %>% 
  select(stop_islam:muslim_rad)

islam_pmm_imp_sum <- sapply(islam_pmm_imp_final, 
  plyr::each(mean, median, sd, var, min, max)) %>%
  t() %>%
  as.data.frame() %>%
  mutate(method = "Predictive mean matching")
```



```{r}
library(tidyr)
library(mice)
library(ggplot2)
library(ggthemes)

### help function
norm_dens <- function(dat) {
    mi <- min(dat$value)
    ma <- max(dat$value)
    grid <- seq(mi, ma, by = .1)
    return(data.frame( 
    predicted = grid,
    density = dnorm(grid, mean(dat$value), sd(dat$value))
    ))
}

### transform function
# imp_object <- islam_lm_imp
# dest_vars <- dv

vis_imputation <- function(imp_object, dest_vars, means = F){
  # finalize imp data
  imp_stacked <- mice::complete(imp_object,
                         "long",
                         include = F)
  # select main variables
  imp_stacked <- imp_stacked[, c(".imp", ".id", dest_vars)]
  
  # imputation step
  #levels(imp_stacked$.imp) <- paste("Imputation", 1:5)
  
  # NA matrix + melt
  na_matrix <- as.data.frame(is.na(imp_object$data))
  na_matrix$.id <- unique(imp_stacked$.id)
  na_matrix$.imp <- imp_stacked$.imp[1:nrow(na_matrix)]

  na_matrix_long <- na_matrix[, c(".imp", ".id", dest_vars)] %>% 
    gather(var, miss, -.imp, -.id)

  # imp data melt 
  imp_dat_long <- imp_stacked %>% 
    gather(var, value, -.imp, -.id)

  # merge data
  names(imp_dat_long)
  gg_final <- cbind(imp_dat_long, miss = na_matrix_long$miss)

  # visualize long format
  normaldens <- gg_final %>%
  group_by(var, miss) %>%
  do(norm_dens(.))

  gg1 <- ggplot(data = gg_final, aes(x = value, fill = miss)) + 
    #geom_density(binwidth = 1, alpha = .5) +
    geom_histogram(aes(y = ..density..), 
                   position = position_dodge(), 
                   binwidth = 1, 
                   alpha = .5) +
    geom_line(data = normaldens, 
              aes(x = predicted, y = density, colour = miss)) + 
    scale_colour_manual(name = "", 
                      values = c("#3A5FCD", "#EE0000"),
                      labels = c("observed", "imputed")) +
    scale_fill_manual(name = "", 
                      values = c("#3A5FCD", "#EE0000"),
                      labels = c("observed", "imputed")) +
    facet_wrap(~ var, scales = "free", ncol = 3) +
    theme_classic() +
    theme(legend.position = "bottom") + 
    labs(x = "", y = "") 
  
  if(means){
  # create means
      m <- gg_final %>%
      group_by(var, miss) %>%
      summarise(m1 = mean(value))
      
      gg1 + geom_vline(data = m, aes(xintercept = m$m1, colour = miss))
  } else { 
    gg1
  }
}
```



## Häufigkeitsverteilungen für Nationalbewusstsein

```{r, results = "asis", fig.height = 8}
impute_vis <- function(data, methods, labels){
  for(i in seq_along(methods)){
    nation <- dat %>% 
      select(respid:age, isced97, ger_citiz:born_ger)

    # default methods
    method <- c(rep("", 4) ,rep("pmm", 11))
    # avoid negative imputed values
    post <- c(rep("", 4) ,rep("imp[[j]][,i]<-squeeze(imp[[j]][,i],c(0,7))", 11))

    ### run linear imputation
    nation_imp <- mice(nation,
                  method = methods[i],
                  post = post,
                  seed = 1234, 
                  printFlag = F)
    
    dv <- colnames(select(nation_imp$dat, ger_citiz:born_ger))
    gg1 <- vis_imputation(nation_imp, dest_vars = dv, means = F) +
      ggtitle(paste0("Item-Batterie Nationalbewusstsein: ", labels[i]))
    print(gg1)
  }
  
}

methods <- c("mean",
             "norm.predict",
             "norm.boot",
             "norm.nob",
             "norm",
             "pmm")

labels <- c("Unconditional mean imputation",
            "Linear regression, predicted values",
            "Linear regression using bootstrap",
            "Linear regression ignoring model error",
            "Bayesian linear regression",
            "Predictive mean matching")

impute_vis(dat, methods, labels)
```


## Deskriptive Statistik für unterschiedlich imputierte Variablen Nationalbewusstsein

```{r, results = "asis", comment = F, eval = T}
nation_all <- rbind(nation_ld_sum,
                   nation_mean1_sum,
                   nation_lm_imp_sum,
                   nation_pmm_imp_sum)

k_data <- 4
nation_all <- cbind(vars = rep(rownames(nation_ld_sum), k_data), nation_all) 

nation_all <- nation_all %>% arrange(vars)
nation_all <- nation_all %>% select(vars, method, mean:var)

nation_all$vars[duplicated(nation_all$vars)] <- ''

library(stargazer)
stargazer(nation_all, 
          summary = F,
          header = F,
          digits = 2, 
          rownames = F, 
          type = "latex",
          #type = "html",
          style = "ajps",
          title = "Alle")
```


## Häufigkeitsverteilungen für Islamophobie

```{r, results = "asis", fig.height = 4, eval = F}
dv <- colnames(select(islam_mean2_imp$dat, stop_islam:muslim_rad))
vis_imputation(imp_object = islam_mean2_imp, dest_vars = dv, means = F) +
  ggtitle("Item-Batterie Islamophobie: Mean_Imputation")
```


```{r, results = "asis", fig.height = 4}
### Islamophobie
# select data for imp 

impute_vis <- function(data, methods, labels){
  for(i in seq_along(methods)){
    islam <- data %>% 
    select(respid:age, isced97, stop_islam:muslim_rad)
    
    # default methods
    method <- c(rep("", 4) ,rep("pmm", 6))
    # avoid negative imputed values
    post <- c(rep("", 4) ,
              rep("imp[[j]][,i]<-squeeze(imp[[j]][,i],c(0,7))", 6))
    
    ### run linear imputation
    islam_imp <- mice(islam,
                  method = methods[i],
                  post = post,
                  seed = 1234, 
                  printFlag = F)
    
    dv <- colnames(select(islam_imp$dat, stop_islam:muslim_rad))
    gg1 <- vis_imputation(islam_imp, dest_vars = dv, means = F) +
      ggtitle(paste0("Item-Batterie Islamophobie: ", labels[i]))
    print(gg1)
  }
  
}

methods <- c("mean",
             "norm.predict",
             "norm.boot",
             "norm.nob",
             "norm",
             "pmm")

labels <- c("Unconditional mean imputation",
            "Linear regression, predicted values",
            "Linear regression using bootstrap",
            "Linear regression ignoring model error",
            "Bayesian linear regression",
            "Predictive mean matching")

impute_vis(dat, methods, labels)
```



The function `mice.impute.pmm()` implements predictive mean matching (Little 1988), a general purpose semi-parametric imputation method. Its main virtues are that imputations are restricted to the observed values and that it can preserve non-linear relations even if the structural part of the imputation model is wrong. A disadvantage is that it may fail to produce enough between-imputation variability if the number of predictors is small.


## Deskriptive Statistik für unterschiedlich imputierte Variablen Islamophobie

```{r, results = "asis", comment = F, eval = T}
islam_all <- rbind(islam_ld_sum,
                   islam_mean1_sum,
                   islam_lm_imp_sum,
                   islam_pmm_imp_sum)

k_data <- 4
islam_all <- cbind(vars = rep(rownames(islam_ld_sum), k_data),
                   islam_all) 

islam_all <- islam_all %>% arrange(vars)
islam_all <- islam_all %>% select(vars, method, mean:var)

islam_all$vars[duplicated(islam_all$vars)] <- ''

library(stargazer)
stargazer(islam_all, 
          summary = F,
          header = F,
          digits = 2, 
          rownames = F, 
          type = "latex",
          #type = "html",
          style = "ajps",
          title = "Alle")
```


```{r, result = "asis", eval = F}
## Load VIM package for Visualization and Imputation of Missing Values
library(VIM)

## in number
aggr(dat, prop = F, numbers = T)

## in proportions
aggr(dat, prop = T, numbers = T)

## Matrix plot. Red for missing values, Darker values are high values.
# matrixplot(dat, interactive = F, sortby = NULL)

## Margin plot. Red dots have at least one missing. No observation with two missing values here.
# marginplot(sleep[,c("Gest","Dream")])
```


```{r, result = "asis", eval = F}
scattmatrixMiss(dat[,7:16], interactive = F)
scattmatrixMiss(dat[,17:22], interactive = F)
library(mice)
# md.pattern(dat)
```


```{r}
## Create data frame indicating missingness by 1
x <- as.data.frame(abs(is.na(dat)))
## Select columns with some (but not all) missing values
y <- x[,sapply(x, sd) > 0]

## Create a correlation matrix: Variables missing together have high correlation
cor_tab <- round(cor(y), 2)
cor_tab[lower.tri(cor_tab)]<- ''
# cor_tab
```

\blandscape
```{r xtable, results="asis", eval = F}
library(xtable)
print(xtable(cor_tab, caption="Pearson's correlation matrix of missing values"),
      rotate.colnames = T, size = "small", comment = F, missing = "?")
```
\elandscape


